### 2009-10-20 latex/paper1/contents.tex

%One of the most obvious ways to reduce the time needed to develop or maintain a
%software system is to assign more developers to it. When combined with poor
%management, though, adding developers may in fact raise the development time,
%due to the increasing communication costs between developers \cite{Brooks1995}.
%
%The communication costs are affected by the assignment of developers to
%implementation components of a system.
%%The assignment of developers to implementation components of a system affects
%%the amount of communication needed between developers. 
%If two development teams are responsible for maintaining components that are
%highly interdependent, they need to coordinate their work often in order avoid
%code duplication and breaking each other's component funcionality.

Therefore, to optimize the development of a system, a manager can decompose the
system into modules, or groups of components, so that there are as few
inter-module dependencies as possible. % (i.e., the modules are weakly couple.).
%Each module is, then, threated as an unit of work assignment \cite{Parnas1972}:
Then, each module is assigned to one team and, because of the little inter-module
dependency, the communication between teams is reduced, therefore speeding up
the development.

@@

In our research we propose to evaluate software clustering algorithms by
applying them to synthetic, i.e., computer generated, component dependency
networks with built-in reference decompositions. This approach enables one to
test algorithms with arbitrarily large samples in a controlled manner.

A desirable property of synthetic networks is that they resemble their natural
counterparts. In this paper, we present three network models and show
empirically that all three models are capable of synthesizing networks that
resemble software networks. We consider object-oriented systems with classes as
components.

@@

Network theory studies general properties of many types of networks by using
statistical analysis. In the last decade, remarkable similarities have been
found in networks arising from domains as diverse as sociology, biology,
technology, and linguistics. It has been shown that, in these networks, the
distribution of vertex degrees is a power law, i.e., the number of vertices
connected to $k$ edges is proportional to $k^{-\gamma}$, where $\gamma$ is a
positive constant. These networks are called scale-free networks
\cite{Barabasi1999}.

Network theory has been applied to class dependency networks, represented as
directed unweighted graphs, and it was shown that they are also scale-free
networks \cite{Myers2003,Valverde2002}. Thus, the study of the structure of
object-oriented software systems can benefit from the vast literature available
about scale-free networks.

In particular, many scale-free network models were proposed. These models are
algorithms that were proven, either formally or empirically, to synthesize
networks that are scale-free. In this section we present three models that
synthesize scale-free networks with built-in modular decompositions: BCR+, CGW,
and LF. The networks are represented as directed unweighted graphs with modules,
which makes them suitable for testing software clustering algorithms.

@@

First the model creates an initial network in which each module contains exactly
one vertex. Then all allowed external edges are added. Finally, the networks is
modified by sucessive choices between three operations that add vertices or
edges to the network, until it reaches $n$ vertices. Each time, the probability
of choosing the $i$-th operation is $p_i$.

@@

The LF model \cite{Lancichinetti2009} is a very flexible model that can
generated weighted directed networks with overlapping modules, that is, in which
a vertex can belong to more than one module. Unlike the previous models, this is
not a growth model: all vertices are generated at once and then the edges are
added.

There is also a special version of the model in which the edges weights are
discarded and the modules do not overlap. We used the original implementation of
this version, available at
\url{http://santo.fortunato.googlepages.com/inthepress2}. It accepts the
following parameters:

@@

We already know that the models synthesize networks that, just like software
networks, are scale-free. This single property, though, isn't enough to prove
the hypothesis, since many known scale-free networks are not software networks
(e.g., biologic networks). Therefore, the similarity metric should be able to
distinguish between software and non-software networks.

In this section we present a similarity metric and validate the metric by
appplying it to a data set containing both software and non-software networks.

@@

To support the evaluation of the metric, we have collected 131 networks. The
networks are described in detail in the Appendix.

\subsubsection{Software networks} We have collected 65 software systems written in
Java, with size ranging from 111 to 35,363 classes. Java was chosen for being a
popular programming language in which many open source systems have been
written. The software networks, representing dependencies between classes, were
extracted with the tool Dependency Finder\footnote{Available at
\url{http://depfind.sf.net}.}.

@@

We chose $p + q > 0$ because otherwise no node would be added after the
creation of the initial network. We also avoided large values for $\mu$ in
order to ignore networks with strongly coupled modules.

@@

Notice that $e_i$ has no effect when $p_i = 0$; in this case $e_i$ was just set
to zero.

@@

\begin{table}
\caption{Rules for predicting the classification of a synthetic network. S
stands for software-like and N stands for non software like.}
\centering
\begin{tabular}{|l|l|l|}
\hline
Model & Rule & Accuracy \\
\hline 
\hline
\multirow{2}{*}{BCR+}
     & $p_1 \ge 0.7 \Rightarrow S$ & \multirow{2}{*}{82.4\%}  \\ 
     & $p_1 < 0.7 \Rightarrow N$ & \\ 
\hline
\multirow{2}{*}{CGW}
     & $p_1 \ge 0.5 \Rightarrow S$ & \multirow{2}{*}{82.3\%} \\  
     & $p_1 < 0.5 \Rightarrow N$ & \\  
\hline
\multirow{2}{*}{LF}   
     & $\gamma \ge 2.44 \Rightarrow S$ & \multirow{2}{*}{78.9\%} \\ 
     & $\gamma < 2.44 \Rightarrow N$ & \\ 
\hline
\end{tabular}
\label{tab:rules}
\end{table}

@@

In a future work, we intend to use synthetic networks in the evaluation of
software clustering algorithms that were previously tested with real networks.
After that we will be able to compare the results obtained by different
approaches.

>>>>

Therefore, to optimize the development of a system, a manager can decompose the
system into modules, or groups of components, so that there are as few
inter-module dependencies as possible. % (i.e., the modules are weakly coupled).
%Each module is, then, threated as an unit of work assignment \cite{Parnas1972}:
Then, each module is assigned to one team and, because of the little inter-module
dependency, the communication between teams is reduced, therefore speeding up
the development.

%********** 
%
%Large software systems need to be decomposed in smaller parts in order to be
%effectively maintained by groups of developers working concurrently. If a system
%is poorly documented, 
%
%**********
%
%Even if a
%system is poorly documented, a software manager should be able to decompose it into 
%modules with few dependencies between them, so 
%
%find relatively
%independent modules in it, so each development team 
%
% A software
%manager should be
%**********
%
% The task of
%finding cohesive modules in the source code
%
%Although
%dependency analysis tools can aid in the task of 
%Dependency
%analysis

@@

An alternative approach is to run the algorithms on synthetic, i.e., computer-
generated, component dependency networks with built-in reference decompositions.
This approach enables one to test algorithms with arbitrarily large samples in a
controlled manner. Nevertheless, this approach is often overlooked, mainly
because naïve random network models synthesize networks that are very dissimilar
from component dependency networks.

In this paper, we present three recently proposed network models and show
empirically that they are capable of synthesizing networks that resemble
software component dependency networks. We limit our analysis to networks of
static dependencies between classes in object-oriented systems written in Java.

@@

Network theory studies general properties of many types of networks by using
statistical analysis. In the last decade, remarkable similarities have been
found in networks arising from domains as diverse as sociology, biology,
technology, and linguistics. It has been shown that, in many networks, the
distribution of vertex degrees is a power law, i.e., the number of vertices
connected to $k$ edges is proportional to $k^{-\gamma}$, where $\gamma$ is a
positive constant. Networks that share this property are called scale-free
networks \cite{Barabasi1999}.

Network theory has been applied to class dependency networks, represented as
directed unweighted graphs, and it was shown that they are also scale-free
\cite{Myers2003,Valverde2002}. Thus, the study of the structure of
object-oriented software systems can benefit from the vast literature available
about scale-free networks.

Scale-free network models are algorithms that were proven, either formally or
empirically, to synthesize networks that are scale-free. In this section we
present three models that synthesize scale-free networks with built-in modular
decompositions: BCR+, CGW, and LF. The networks are represented as directed
unweighted graphs with modules, which makes them suitable for testing software
clustering algorithms.

@@

First the model creates a network in which each module contains exactly one
vertex. Then all allowed external edges are added. Finally, the networks is
modified by sucessive choices between three operations that add vertices or
edges to the network, until it reaches $n$ vertices. Each time, the probability
of choosing the $i$-th operation is $p_i$.

@@

The LF model \cite{Lancichinetti2009} is a very flexible model that can
generate weighted directed networks with overlapping modules, that is, in which
a vertex can belong to more than one module. Unlike the previous models, this is
not a growth model: all vertices are generated at once and then the edges are
added.

There is also a special version of the model in which the edge weights are
discarded and the modules do not overlap. We used the original implementation of
this version, available at
\url{http://santo.fortunato.googlepages.com/inthepress2}. It accepts the
following parameters:

@@

We already know that the models synthesize networks that, just like software
networks, are scale-free. This single property, though, is insufficient to prove
the hypothesis, since many known scale-free networks are not software networks
(e.g., biologic networks). Therefore, the similarity metric should be able to
distinguish between software and non-software networks.

In this section we present a similarity metric and validate it by appplying it
to a data set containing both software and non-software networks.

@@

To support the evaluation of the metric, we have collected 131 networks from
many different domains. The networks are described in detail in the Appendix.

\subsubsection{Software networks} We have collected 65 software systems written
in Java, with size ranging from 111 to 35,363 classes. Java was chosen for being
a popular programming language in which many open source systems have been
written. The software networks, representing static dependencies between
classes, were extracted with the tool Dependency Finder\footnote{Available at
\url{http://depfind.sf.net}.}.

@@

We chose $p + q > 0$ because otherwise no node would be added to the initial
network. We also avoided large values for $\mu$ in order to ignore networks with
strongly coupled modules.

@@

Note that $e_i$ has no effect when $p_i = 0$; in this case $e_i$ was just set to
zero.

@@

\begin{table}
\caption{Rules for predicting the classification of a synthetic network. S
stands for software-like and N stands for non software like.}
\centering
\begin{tabular}{|l|l|l|}
\hline
Model & Rule & Accuracy \\
\hline 
\hline
\multirow{2}{*}{BCR+}
     & $p_1 \ge 0.7 \Rightarrow S$ & \multirow{2}{*}{82.4\%}  \\ 
     & $p_1 < 0.7 \Rightarrow N$ & \\ 
\hline
\multirow{2}{*}{CGW}
     & $p_1 \ge 0.5 \Rightarrow S$ & \multirow{2}{*}{82.3\%} \\  
     & $p_1 < 0.5 \Rightarrow N$ & \\  
\hline
\multirow{2}{*}{LF}   
     & $\gamma < 2.44 \Rightarrow S$ & \multirow{2}{*}{78.9\%} \\ 
     & $\gamma \ge 2.44 \Rightarrow N$ & \\ 
\hline
\end{tabular}
\label{tab:rules}
\end{table}

@@

In a future work, we intend to use synthetic networks in the evaluation of
software clustering algorithms that were previously tested with real networks.
After that we will be able to compare the results obtained by the two
approaches.
